### 2020-11-04 latex/acrobot.tex
Updated the acrobot stuff again, still on oscillations.

When \(I = 0\), the constrained acrobot behaves like a single
pendulum with masses at a distance \(l\) and \(2l\) from the pivot 
(Figure \ref{fig:acrobot-I0}) whose energy
\[
    E(q_u,p_u) = \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
    ,
\]
is conserved.
Level sets of \(E\) are ellipses when \(E \leq E(\pi,0)\), which we call
``oscillations"; and they are open curves when \(E > E(\pi,0)\), which we call
``rotations". 
Examples of these can be seen in Figure \ref{fig:pendulum-level-sets}.

@@

\subsection*{Perturbation Analysis for Oscillations}
An oscillation of the nominal pendulum (when \(I = 0\)) is uniquely determined
by its intersection \(\mu\) with the \(q_u\)-axis, as in Figure
\ref{fig:mu-intersection}. Since \(\mu\) is to be a pseudo-radius and
\(q_u \in ]-\pi,\pi]\), we have \(\mu \in ]0,\pi]\). 
The corresponding energy level set is \(E(q_u,p_u) = 3mgl(1- \cos(\mu))\), which
gives the relationship
\begin{equation}\label{eqn:oscillation-pu2}
    p_u^2 = 30m^2gl^3(\cos(q_u) - \cos(\mu))
    .
\end{equation}

\begin{figure}
    \centering
    % TODO: Add a figure with one oscillation whose we mark the distance mu from
    % the origin to the intersection on the q-axis
    %\includestandalone[]{images/mu_intersection}
    \caption{The pseudo-radius \(\mu\) for oscillations is within \(]0,\pi]\).}
    \label{fig:mu-intersection}
\end{figure}

On this level set, \(q_u\) ranges between \([-\mu,\mu]\) and can be uniquely
parameterized by \(q_u = \mu \cos(\alpha)\), where \(\alpha\) is our
pseudo-angle.
Substituting this into (\ref{eqn:oscillation-pu2}), we get
\[
    p_u^2 = 30m^2gl^3(\cos(\mu \cos(\alpha)) - \cos(\mu))
    .
\]
We want to find \(p_u\) as a function of \((\alpha,\mu)\); noting that we can
determine the sign of \(p_u\) from the sign of \(\sin(\alpha)\), we get the
(clockwise) parameterization
\begin{equation}\label{eqn:oscillation-pu}
    p_u = -\sign{\sin(\alpha)} \sqrt{30m^2gl^3 \left(\cos(\mu c_\alpha) - c_\mu\right)}
    ,
\end{equation}
which is, in fact, continuously differentiable.

We have thus found a mapping \((\alpha,\mu) \mapsto (q_u,p_u)\) for which we
need the inverse mapping.
Notice from (\ref{eqn:oscillation-pu2}) that
\[
    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
    .
\]
Since \(\mu \in ]0,\pi]\), we can uniquely express \(\mu\) by
\begin{equation}\label{eqn:mu-qupu}
    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
    .
\end{equation}

>>>>

When \(I = 0\), the constrained acrobot behaves like a single
pendulum with masses at a distance \(l\) and \(2l\) from the pivot 
(Figure \ref{fig:acrobot-I0}) whose energy
\begin{equation}\label{eqn:acrobot-nominal-E}
    E(q_u,p_u) = \frac{p_u^2}{10ml^2} + 3(1 - \cos(q_u))
    ,
\end{equation}
is conserved.
Level sets of \(E\) are ellipses when \(E < E(\pi,0)\), which we call
``oscillations"; and they are open curves when \(E > E(\pi,0)\), which we call
``rotations". 
Examples of these can be seen in Figure \ref{fig:pendulum-level-sets}.

@@

\subsection*{Perturbation Analysis for Oscillations}
The nominal pendulum (\ie when \(I = 0\)) oscillates whenever the
nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
Hence, the domain of oscillations on \(\Gamma\) is given by
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
and \(\mu > 0\) is a pseudo-radius.
Since the energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

\begin{figure}
    \centering
    % TODO: Add a figure with one oscillation mark at a distance mu from
    % the origin to the intersection on the q-axis. We also draw the boundary of
    % D and shade the interior
    %\includestandalone[]{images/mu_intersection}
    \caption{The domain \(D\) where a pendulum oscillates.
    The pseudo-radius coordinate \(\mu\) corresponds to the
    intersection of an orbit of oscillation with the \(q_u\)-axis.}
    \label{fig:mu-intersection}
\end{figure}

The energy level set corresponding to the intersection \((q_u,p_u) = (\mu,0)\)
is 
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = 3mgl(1- \cos(\mu))\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:oscillation-pu2}
    p_u^2 = 30m^2gl^3(\cos(q_u) - \cos(\mu))
    .
\end{equation}
On this level set, \(q_u\) ranges between \([-\mu,\mu]\) and can be uniquely
parameterized by \(q_u = \mu \cos(\alpha)\), where \(\alpha\) is our
pseudo-angle.
Substituting this into (\ref{eqn:oscillation-pu2}), we get
\[
    p_u^2 = 30m^2gl^3(\cos(\mu \cos(\alpha)) - \cos(\mu))
    .
\]
We want to find \(p_u\) as a function of \((\alpha,\mu)\); noting that we can
determine the sign of \(p_u\) from the sign of \(\sin(\alpha)\), we get the
(clockwise) parameterization
\begin{equation}\label{eqn:oscillation-pu}
    p_u = -\sign{\sin(\alpha)} \sqrt{30m^2gl^3 \left(\cos(\mu c_\alpha) - c_\mu\right)}
    ,
\end{equation}
which is, in fact, smooth.

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\).
We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
Notice from (\ref{eqn:oscillation-pu2}) that
\[
    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
    .
\]
Since \(\mu \in ]0,\pi]\), we can uniquely express \(\mu\) by
\begin{equation}\label{eqn:mu-qupu}
    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
    .
\end{equation}
Next we need to find \(\alpha\). 
Recall that 
\[
    \cos(\alpha) = \frac{q_u}{\mu}
    ,
\] 
which means 
\[
    \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
    .
\]
Notice from (\ref{eqn:oscillation-pu}) that 
\(\sign{\sin(\alpha)} = -\sign{p_u}\).
We use these facts to deduce that
\begin{equation}\label{eqn:alpha-qupu}
    \alpha = \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
    ,
\end{equation}
and thus have our transformation \(T(q_u,p_u)\) into
\((\alpha,\mu)\)-coordinates on \(D\).

The acrobot's constrained dynamics in
\((\alpha,\mu)\)-coordinates can be computed by evaluating
\[
    \left.\begin{cases}
        \dot{\mu} = 
        \begin{bmatrix} \pdiff{\mu}{q_u} & \pdiff{\mu}{p_u} \end{bmatrix}
        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
        , \\
        \dot{\alpha} = \begin{bmatrix} \pdiff{\alpha}{q_u} & \pdiff{\alpha}{p_u} \end{bmatrix}
        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
        .
    \end{cases} \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
\]
These dynamics are incredibly complicated and difficult to express, so we will
write them simply as
\[
    \begin{cases}
        \dot{\mu} = f_\mu(\alpha,\mu,I)
        ,\\
        \dot{\alpha} = f_\alpha(\alpha,\mu,I)
        .
    \end{cases}
\]

To find the nominal system we set \(I = 0\); 
MATLAB's symbolic toolbox operations tell us that
\begin{equation}\label{eqn:acrobot-nominal-ma-dynamics}
    \begin{cases}
        \dot{\mu} = 0
        , \\
        \dot{alpha} = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
        .
    \end{cases}
\end{equation}
